---
title: "Ideas and Experiments"
author: "Orrego Chong, Kantelhardt, Kornelius"
date: "2023-05-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
RDeeP_HeLa_Mitosis = read.csv("https://www.dropbox.com/s/ttj1z98mxefq0aj/RDeeP_HeLa_Mitosis.csv?dl=1", header=TRUE, row.names=1, sep = ";")

head(RDeeP_HeLa_Mitosis)
```
```{r}
#Deleting the complete zero rows:
RDeeP_HeLa_Mitosis <- RDeeP_HeLa_Mitosis[-which(rowSums(RDeeP_HeLa_Mitosis)==0),]
```


### Student t-test 

```{r}
# Extract the first three values from the column
subset_values <- RDeeP_HeLa_Mitosis[1, 1:3]

# Perform the t-test
t_test_result <- t.test(subset_values)

# Print the t-test result
print(t_test_result)
```
### Reorder dataset

```{r}
RDeeP_HeLa_Mitosis_Ctrl = cbind(RDeeP_HeLa_Mitosis[,1:3],RDeeP_HeLa_Mitosis[,7:9],RDeeP_HeLa_Mitosis[,13:15],RDeeP_HeLa_Mitosis[,19:21],RDeeP_HeLa_Mitosis[,25:27],RDeeP_HeLa_Mitosis[,31:33],RDeeP_HeLa_Mitosis[,37:39],RDeeP_HeLa_Mitosis[,43:45],RDeeP_HeLa_Mitosis[,49:51],RDeeP_HeLa_Mitosis[,55:57],RDeeP_HeLa_Mitosis[,61:63],RDeeP_HeLa_Mitosis[,67:69],RDeeP_HeLa_Mitosis[,73:75],RDeeP_HeLa_Mitosis[,79:81],RDeeP_HeLa_Mitosis[,85:87],RDeeP_HeLa_Mitosis[,91:93],RDeeP_HeLa_Mitosis[,97:99],RDeeP_HeLa_Mitosis[,103:105],RDeeP_HeLa_Mitosis[,109:111],RDeeP_HeLa_Mitosis[,115:117],RDeeP_HeLa_Mitosis[,121:123],RDeeP_HeLa_Mitosis[,127:129],RDeeP_HeLa_Mitosis[,133:135],RDeeP_HeLa_Mitosis[,139:141],RDeeP_HeLa_Mitosis[,145:147])

#For RNase it is easier, because we now can just take all the row which were not taken in the first table. 
RDeeP_HeLa_Mitosis_RNase=RDeeP_HeLa_Mitosis[,-which(names(RDeeP_HeLa_Mitosis) %in% colnames(RDeeP_HeLa_Mitosis_Ctrl))]

head(RDeeP_HeLa_Mitosis_RNase)
```


## 2. Scale and Reduce Dataset

#### 2.1 Normalization (scaling, or z-transformation?)

Old complicated version for finding mean:

1. Create an empty dataframe to store the mean values
mean_values <- data.frame(matrix(ncol = ncol(RDeeP_HeLa_Mitosis_Ctrl)/3, nrow = nrow(RDeeP_HeLa_Mitosis_Ctrl)))

2.Outer loop: Iterate over every row (protein) in the original table
for (i in 1:nrow(RDeeP_HeLa_Mitosis_Ctrl)) {
  # Inner loop: Iterate over every three columns at a time
  for (j in seq(1, ncol(RDeeP_HeLa_Mitosis_Ctrl), by = 3)) {
    # Select the three columns for the current iteration
    selected_columns <- RDeeP_HeLa_Mitosis_Ctrl[i, j:(j + 2)]
    
    # Calculate the mean of the selected columns
    mean_value <- mean(selected_columns, na.rm = TRUE)
    
    # Assign the mean value to the corresponding cell in the mean_values dataframe
    mean_values[i, (j + 2) %/% 3] <- mean_value
  }
}

NEW VERSION:

```{r}
#First finding the mean of all three replicates for every fraction separately:
Mitosis_Ctrl_mean = as.data.frame(
  sapply(seq(0,72,3), function(i) { #outer loop to repeatedly select the three replicates for each fraction
    apply(RDeeP_HeLa_Mitosis_Ctrl[,c(1,2,3)+c(i,i,i)], 1, function(x) {mean(x)})
    #inner loop which calculates the mean for the three replicates, for each row anew. 
    
    }))

#To have logical new column names, it is possible to paste the fraction number and add Ctrl_Mean separately to each new column name.
colnames(Mitosis_Ctrl_mean) = paste("Fraction",seq(1,25,1),"_Ctrl_Mean",sep="")

head(Mitosis_Ctrl_mean)
```

Now trying data scaling to 100:
```{r}
Mitosis_Ctrl_100 <- as.data.frame(sapply(Mitosis_Ctrl_mean, function(i){(i/apply(Mitosis_Ctrl_mean,1,sum))*100}))

head(Mitosis_Ctrl_100)

```
Seems to work? Now all the same for RNase:

```{r}
#First finding mean:
Mitosis_RNase_mean = as.data.frame(
  sapply(seq(0,72,3), function(i) {
    apply(RDeeP_HeLa_Mitosis_Ctrl[,c(1,2,3)+c(i,i,i)], 1, function(x) {mean(x)})
    }))

colnames(Mitosis_RNase_mean) = paste("Fraction",seq(1,25,1),"_RNase_Mean",sep="")

#Now scaling 
Mitosis_RNase_100 <- as.data.frame(sapply(Mitosis_RNase_mean, function(i){(i/apply(Mitosis_RNase_mean,1,sum))*100}))

head(Mitosis_RNase_100)

```


```{r}
#Then maybe applying the z transformation on each row was a whole, Control and RNase separate? Still needs a lot of refining...
#Mitosis_Ctrl_norm <- apply(t(Mitosis_Ctrl_mean),1, function(x)(x - mean(x)) / sd(x))

#transpose is necessary as "apply" is always iterates the command or each column 

head(Mitosis_Ctrl_mean)
```


Min Max scaling?




