---
title: "Topic 3: Proteome-wide Screen for RNA-dependent Proteins; Subtopic 3: HeLa synchronized in Mitosis"
author: "Jimena, Alexandra, Verena"
date: "2023-05-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Topic 3.3: HeLa Cells Synchronized in Mitosis

### Proteome-wide Screen for RNA-dependent Proteins

Load the dataset from the Dropbox

```{r}
RDeeP_HeLa_Mitosis = read.csv("https://www.dropbox.com/s/ttj1z98mxefq0aj/RDeeP_HeLa_Mitosis.csv?dl=1", header=TRUE, row.names=1, sep = ";")
head(RDeeP_HeLa_Mitosis)
```

## 1. Preparing data for analysis (Data Cleanup)

#### 1.1 Data description

```{r}
dim(RDeeP_HeLa_Mitosis)
head(rownames(RDeeP_HeLa_Mitosis))
head(colnames(RDeeP_HeLa_Mitosis))
```

#### 1.2 Finding rows with missing values

Finding which rows are all 0 (and are most likely experimental errors). And removing this rows.

```{r}
which(rowSums(RDeeP_HeLa_Mitosis)==0)

RDeeP_HeLa_Mitosis = RDeeP_HeLa_Mitosis[-which(rowSums(RDeeP_HeLa_Mitosis)==0)]
#Tabelle ohne die Nullzeile wieder in der Variable gespeichert

```

#### 1.3 Handeling other wierd rows

When looking at the dataset we noticed that already the first 7 rows have accumulation of protein in every fraction. We are not sure if we can work with them as they maybe have no clear peak, therefore we first identify them and then maybe delete them later.

```{r}
# Find rows without any zeros
nonzero_rows <- RDeeP_HeLa_Mitosis[rowSums(RDeeP_HeLa_Mitosis == 0) == 0, ]

#The condition rowSums(RDeep_Hela_Mitosis == 0) == 0 checks which rows have a sum of zeros equal to zero, indicating that those rows do *not* contain any zeros.Then we selects these rows without any zeros from our table and store it in a new table.

# Print the rows without any zeros
print(nonzero_rows)
nrow(nonzero_rows) #320 rows do not have zeros


#Possibly deleting these nonzero rows and storing it in the original matrix:
#RDeeP_HeLa_Mitosis <- RDeeP_HeLa_Mitosis[rowSums(RDeeP_HeLa_Mitosis == 0) > 0, ]

```

#### 1.4 Check data format

Here we try to determine data type of our dataset, whether it is numerical, categorical, etc
```{r}


nrow(RDeeP_HeLa_Mitosis)
ncol(RDeeP_HeLa_Mitosis)
sum(apply(RDeeP_HeLa_Mitosis, 1, is.na)) # Summe aller Spalten, die NA Werte haben
sum(apply(RDeeP_HeLa_Mitosis, 2, is.na)) # Summe aller Zeilen, die NA Werte haben
sum(apply(RDeeP_HeLa_Mitosis, 1, is.numeric))
sum(apply(RDeeP_HeLa_Mitosis, 2, is.numeric))

#sum(apply(RDeeP_HeLa_Mitosis, 1, is.numeric)) == nrow(RDeeP_HeLa_Mitosis)

```

This means we have numeric data, aka the data from the mass spectrometry.

```{r}
#Complete DataSet
View(RDeeP_HeLa_Mitosis)
```

#### 1.5 Checking reproducibility with student t-test

```{r}
# Extract the first three values from the column
subset_values <- RDeeP_HeLa_Mitosis[1, 1:3]

# Perform the t-test
t_test_result <- t.test(subset_values)

# Print the t-test result
print(t_test_result)
```

#### 1.6 Checking reproducubility with correlation?
