---
title: "Topic 3: Proteome-wide Screen for RNA-dependent Proteins; Subtopic 3: HeLa synchronized in Mitosis"
author: "Jimena, Alexandra, Verena"
date: "2023-05-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Topic 3.3: HeLa Cells Synchronized in Mitosis


# Proteome-wide Screen for RNA-dependent Proteins

Load the dataset from the Dropbox

```{r}
RDeeP_HeLa_Mitosis = read.csv("https://www.dropbox.com/s/ttj1z98mxefq0aj/RDeeP_HeLa_Mitosis.csv?dl=1", header=TRUE, row.names=1, sep = ";")
head(RDeeP_HeLa_Mitosis)
```

## 1. Preparing data for analysis

#### 1.1 Data description

```{r}
dim(RDeeP_HeLa_Mitosis)
head(rownames(RDeeP_HeLa_Mitosis))
head(colnames(RDeeP_HeLa_Mitosis))
```

#### 1.2 Finding rows with missing values

Finding which rows are all 0 (and are most likely experimental errors). And removing this rows.

```{r}
which(rowSums(RDeeP_HeLa_Mitosis)==0)
```

#### 1.3 Check data format

```{r}
sum(apply(RDeeP_HeLa_Mitosis, 1, is.numeric)) == nrow(RDeeP_HeLa_Mitosis)
```

This means we have numeric data, aka the data from the mass spectrometry.

```{r}
#Complete DataSet
View(RDeeP_HeLa_Mitosis)
```

